---
title: "pa_3"
output: github_document
---

# Programming Assignment 3 

## Answers to the questions

a. Examine the portion of the script you see below. In your own words what does this section do and why does it work? Demonstrate that you understand the code….

**This part of the code calculates the time landmarks of the vowels. The first two lines get the values of the onset and the offset of the vowels. The third line calculates the duration of the vowels, which is the difference between the offset and the onset. The next five lines calculate the values of the 5 time landmarks by adding the vowel onset value to the product of the vowel duration multiplied by the percentage that corresponds to the time landmark.**

b. In a few short sentences describe the general outline of the script, what the purpose is, and how it acheives this purpose (hint: focus on the section dividers and the comments)

**The code reads the .wav file and the textgrid, obtains information about the vowels, and saves the results in a .csv file. The code is divided in three parts:**


**The first part of the code sets the folder in which the generated .csv file will be stored, chooses a name for the .csv file, deletes the current output file, and creates the file and the headers for the columns in the .csv file.**


**The second part sets up the loop. The code sets the path to the folder where the stimuli files are stored, gets the .wav file, creates strings for .wav file, reads the .wav file and the textgrid,  checks the intervals in the textgrid, sets default values, and adds a string variable for a personal id.**


**The third part of the code runs the loop. It gets the labels of the intervals, gets the onset and offset of the intervals, gets the item and vowel labels, gets the language label, calculates the vowel onset and offset as well as the duration of the vowels,  calculates the time landmarks, gets formants, calculates spectral centroids and trajectory length, prints the results, saves the results to the .csv file, ends the loop, and removes all the objects from the Praat menu.**

c. In a few short sentences describe how the segmenting procedure you used this week differs from that used in pa_2. What are the advantages and disadvantages?

**In pa_2, we had to label the items in the master .wav file and segment the master .wav file into several .wav files (one file per item). Then, we had to manually create a textgrid for each segmented .wav file and added vowel labels. In the end, we ended up with several segmented .wav files and their corresponding textgrids. The loop accessed each .wav file and the corresponding textrgrid to extract information.**


**In pa_3, we did not have to segment the master .wav file into several .wav files. Instead, we added all the necessary labels to the master .wav file and used the loop to extract all the necessary information directly from the master .wav file.**


**The advantage of using the segmenting procedure of pa_3 is that it saves time and labor since we do not have to create many .wav files and textgrids and manually manipulate them. Since there are more steps to follow in the procedure of pa_2, there may be more room for errors. However, the disadvantage is that we have all the items in the same .wav file, which might cause difficulties depending on how we intend to analyze the stimuli. Both procedures could be useful depending on the needs of our study.**

## Load libraries and read the data in the csv file

```{r}
library(tidyverse)
read_csv("data/vowel_data.csv")
```

## Assign the data to an object df

```{r}
df = read_csv("data/vowel_data.csv")
```

## Descriptive Statistics

### 1. Calculate average F1 centroid and standard deviation

```{r}
mean(df$f1_cent)
sd(df$f1_cent)
```

### 2. Calculate average F2 centroid and standard deviation

```{r}
mean(df$f2_cent)
sd(df$f2_cent)
```

### 3. Calculate average trajectory length and standard deviation

```{r}
mean(df$tl)
sd(df$tl)
```


## Plots

### 1. Trajectory length as a function of vowel and language

```{r}
ggplot(df, aes(fill = language, y = tl, x = vowel)) + 
    geom_bar(position="dodge", stat="identity")
```

### 2. F1 as a function of vowel and language

```{r}
ggplot(df, aes(fill = language, y = f1_cent, x = vowel)) + 
    geom_bar(position="dodge", stat="identity")
```


### 3. F2 as a function of vowel and language

```{r}
ggplot(df, aes(fill = language, y = f2_cent, x = vowel)) + 
    geom_bar(position="dodge", stat="identity")
```

## FREEBIE

```{r, plot5, fig.retina=2, fig.height=5, out.width="100%", eval=T}
# Plot both vowel spaces together ---------------------------------------------

#
# This assumes your dataframe is called 'vowels'
# Before it will work you need to change 'eval=F' to 'eval=T'
#

vowel_means <- df %>% 
  group_by(vowel, language) %>% 
  summarize(f1_cent = mean(f1_cent), f2_cent = mean(f2_cent)) %>% 
  ungroup() %>% 
  mutate(order = case_when(vowel == "i" ~ 1, vowel == "a" ~ 2, TRUE ~ 3), 
         vowel = forcats::fct_reorder2(vowel, vowel, order)) %>% 
  arrange(order)

df %>% 
  mutate(vowel = forcats::fct_relevel(vowel, "u", "a", "i")) %>% 
  ggplot(., aes(x = f2_cent, y = f1_cent, color = language, label = vowel)) + 
    geom_text(size = 3.5, alpha = 0.6, show.legend = F) + 
    geom_path(data = vowel_means, aes(group = language, lty = language), 
              color = "grey") + 
    geom_text(data = vowel_means, show.legend = F, size = 7) + 
    scale_y_reverse() + 
    scale_x_reverse() + 
    scale_color_brewer(palette = "Set1") + 
    labs(title = "Vowel space comparison", 
         subtitle = "Spectral centroids of English/Spanish cardinal vowels", 
         y = "F1 (hz)", x = "F2 (hz)") + 
    theme_minimal(base_size = 16)

# -----------------------------------------------------------------------------
```
